\begin{acknowledgments}
  
I would like to thank my supervisor Dr. Jia Yuan Yu for accepting me into his lab and giving me the opportunity to further develop my academic career with this Master's thesis. I have become a much better researcher thanks to him and I am very grateful for his mentorship. Thank you to everyone in our research lab who read over the drafts and provided valuable feedback along the way.
  
Several fellow researchers have been very helpful in writing this thesis. First and foremost, I would like to express the utmost gratitude to Damien Garreau for inspiring a lot of the work in this thesis. Without his previous work, his phone calls, his emails and his pragmatic ideas, this thesis would not be what it is today. Thank you very much Damien. I would also like to thank Thomas Flynn, Nicholas Keriven, Dougal Sutherland, Iacopo Poli and Arthur Gretton, who all took time of out their schedules to answer my questions and inform particular directions of this thesis. They're responses were super encouraging and welcoming to a newbie like myself.
  
I would also like to thank all my colleagues at DRW who have helped me learn about the finance industry over the past three years. I am especially grateful to Yves, Neil and Laura, who have been very patient with me and have taught me a lot about understanding financial markets, writing clean code and building practical, statistical models.
  
Lastly, I would like to thank my fiancée, Tanya, who initially pushed me to go back to school while I was still young and had the opportunity. 
\end{acknowledgments}
